Visualizing demographic evolution using geographically inconsistent census data

### 2019-08-30 paper/intro.tex
getting ready to send

\unsure{CHECK blue text below}
\revision{1.1}{In some cases, harmonisation might seem methodologically
adequate. If a study is focused on a small subregion that is fixed in time, limiting
and modifying the data to fit that specific region of interest sounds not only
reasonable, but recommended, even if it comes at the cost of producing artificial data points. However, this view is limited in that it ignores interactions beyond the fixed
borders: each different portion of the area of interest may belong to larger
homogeneous processes beyond these arbitrarily defined borders. This effect,
along with the creation of artificial points, is caused by the geometry mismatch
between the data and the region of interest. Instead of forcing one set of
borders into another, our method sidesteps this issue by changing how the data
is represented and interpreted. It therefore allows researchers to examine changes in both the composition and extent of local areas.

@@

To allow a proper examination of our method and its results, we built an online
interactive system using this representation. It enables users to visualise,
interpret, and explore trajectories of neighbourhood change. This interface
helps validate our method, by allowing it to be compared to existing and future
methods. Further, it is a significant contribution to the research community: it
provides a vehicle for quickly and easily grasping complex long-term changes,
experimenting with different parameters to interactively learn from data, and
making neighbourhood change research publicly transparent. The interface thus
responds to increasing concerns about reproducibility and transparency, as well
as ongoing attention to the value of visualisation in scientific research and
communication.

>>>>

\revision{1.1}{In some cases, harmonisation might seem methodologically
adequate. If a study is focused on a small subregion that is fixed in time, limiting
and modifying the data to fit that specific region of interest sounds not only
reasonable, but recommended, even if it comes at the cost of producing artificial data points. However, this view is limited in that it ignores interactions beyond the fixed
borders: each different portion of the area of interest may belong to larger
homogeneous processes beyond these arbitrarily defined borders. This effect,
along with the creation of artificial points, is caused by the geometry mismatch
between the data and the region of interest. Instead of forcing one set of
borders into another, our method sidesteps this issue by changing how the data
is represented and interpreted. It therefore allows researchers to examine changes in both the composition and extent of local areas.

@@

To allow a proper examination of our method and its results, we built a prototype online
interactive system using this representation. It enables users to visualise,
interpret, and explore trajectories of neighbourhood change. This prototype
helps validate our method, by allowing it to be compared to existing and future
methods. Further, it is a significant contribution to the research community: it
provides a vehicle for quickly and easily grasping complex long-term changes,
experimenting with different parameters to interactively learn from data, and
making neighbourhood change research publicly transparent. The interface thus
responds to increasing concerns about reproducibility and transparency, as well
as ongoing attention to the value of visualisation in scientific research and
communication.

### 2019-08-30 paper/method.tex
getting ready to send

\begin{figure*}
    \centering 
    \includegraphics[width=0.975\linewidth]{ui.pdf}
    \caption{Initial interface of our method showing the demographic evolution of Chicago and identifying the objectives of each plot.
        \textbf{(what)}: Cluster overview illustrating the most relevant aspect for each cluster. 
        \textbf{(where)}: Maps illustrating the geographical location, both as an overall summary and for each year.
        \textbf{(how)}: Tabular and visual summmarisation of how the regions classification changed over time.
        \textbf{(who)}: Demographic details for the currently selected data.
        \label{fig:ui}}
\end{figure*}

As illustrated by Figure~\ref{fig:ui}\unsure{CHECK image/ sec 4.5}, our proposed
interface heavily relies on colour to express cluster-related information. We
adopted this convention because colours can be used in all our visual tools in a
coherent manner. However, there is a limit on the number of distinct colours
that can be used. We limited the number of clusters to eight because this was
the largest number of colours that we could reliably and accessibly use, derived
from the 8-class Dark2 set from ColorBrewer~\citep{ColorBrewer}.

>>>>

\begin{figure*}
    \centering 
    \includegraphics[width=0.975\linewidth]{ui.pdf}
    \caption{Initial interface of our method showing the demographic evolution of Chicago and identifying the objectives of each plot.
        \textbf{(what)}: Cluster overview illustrating the most relevant aspect for each cluster. 
        \textbf{(where)}: Maps illustrating the geographical location, both as an overall summary and for each year.
        \textbf{(how)}: Tabular and visual summary of how the regions classification changed over time.
        \textbf{(who)}: Demographic details for the currently selected data.
        \label{fig:ui}}
\end{figure*}

As illustrated by Figure~\ref{fig:ui}, our proposed interface heavily
relies on colour to express cluster-related information. We adopted
this convention because colours can be used in all our visual tools in
a coherent manner. However, there is a limit on the number of distinct
colours that can be used. We limited the number of clusters to eight
because this was the largest number of colours that we could reliably
and accessibly use, derived from the 8-class Dark2 set from
ColorBrewer~\citep{ColorBrewer}.

### 2019-08-30 paper/study.tex
getting ready to send

\unsure{CHECK}\revision{1.12.(1-5)}{These results are meant to demonstrate the
utility of the interface for understanding the evolutionary dynamics of urban
neighbourhoods. They also show the face validity of the results generated by our
novel network-based approach. While our method does not require geographic harmonisation, it does however require matching
the variables over time. This is (currently) a painstaking, manual process, and the current version of the interface remains somewhat limited. For example, we selected a minimal set of
aspects for each country, which are not necessarily similar to each other, or
similar to studies in the current literature. Income moreover is slightly inaccurate,
even though we did correct for official inflation. We grouped the original
ranges into three larger ranges, but they do not match precisely. In addition, the US data labelled "2010" is  not from the decennial census, but from the ACS 2006-2010, somewhat compromising the temporal
stability of the data. Further, the regions selected do not correspond to any
pre-defined regions (metros, cities, census areas), but to somewhat arbitrary
regions defined around a location of interest, making direct comparisons harder.
Taken together, these factors would no doubt require caveats for 
a substantive study using the current prototype version of the interface. Nevertheless, in the present context we find that the interface results are well-aligned
with several other studies, illustrating the robustness of our approach to
imperfect conditions. }

@@

The trajectories plot shows \textit{how} clusters have changed over time.  For
example, it illustrates the process of gentrification, also illustrated in
Figure~\ref{fig:chiWorkflow}, with the purple cluster progressively absorbing
regions from the orange cluster (White). This corroborates results from the
literature reporting that Black neighbourhoods are less likely to
gentrify~\citep{Hwang2014}. The transition matrix provides more details, showing
that around one percent of all regions that became purple at any time
transitioned from the green cluster (Black population), while nine percent were
previouly associated with the orange cluster (White). Next, we select the region
that is gentrified in 2010, by clicking on the corresponding rectangle in the
trajectories plot, updating the information on the maps and the details portion
of the interface.

@@

\unsure{CHECK - maybe add what/how/where/who to figure 5?}

>>>>

\revision{1.12.(1-5)}{These results are meant to demonstrate the
  utility of the interface for understanding the evolutionary dynamics
  of urban neighbourhoods. They also show the face validity of the
  results generated by our novel network-based approach. While our
  method does not require geographic harmonisation, it does however
  require matching the variables over time. This is (currently) a
  painstaking, manual process, and the current version of the
  interface remains somewhat limited. For example, we selected a
  minimal set of aspects for each country, which are not necessarily
  similar to each other, or similar to studies in the current
  literature. Income moreover is slightly inaccurate, even though we
  did correct for official inflation. We grouped the original ranges
  into three larger ranges, but they do not match precisely. In
  addition, the US data labelled "2010" is not from the decennial
  census, but from the ACS 2006-2010, somewhat compromising the
  temporal stability of the data. Further, the regions selected do not
  correspond to any pre-defined regions (metros, cities, census
  areas), but to somewhat arbitrary regions defined around a location
  of interest, making direct comparisons harder.  Taken together,
  these factors would no doubt require caveats for a substantive study
  using the current prototype version of the interface. Nevertheless,
  in the present context we find that the interface results are
  well-aligned with several other studies, illustrating the robustness
  of our approach to imperfect conditions. }

@@

The trajectories plot shows \textit{how} clusters have changed over time.  For
example, it illustrates the process of gentrification, also illustrated in
Figure~\ref{fig:chiWorkflow}, with the purple cluster progressively absorbing
regions from the orange cluster (White). This corroborates results from the
literature reporting that Black neighbourhoods are less likely to
gentrify~\citep{Hwang2014}. The transition matrix provides more details, showing
that around one percent of all regions that became purple at any time
transitioned from the green cluster (Black population), while nine percent were
previously associated with the orange cluster (White). Next, we select the region
that is gentrified in 2010, by clicking on the corresponding rectangle in the
trajectories plot, updating the information on the maps and the details portion
of the interface.

@@
